Interval-Valued Quantum Probabilities: Contextuality and the Born Rule

### 2017-05-10 prop_letter.tex
begin restructure

\title{Contextuality and the Born rule in fuzzy quantum theories}

\author{Andrew J Hanson}
\affiliation{School of Informatics and Computing, Indiana University, Bloomington,
Indiana 47405, USA}

@@

\author{Amr Sabry}
\affiliation{School of Informatics and Computing, Indiana University, Bloomington,
Indiana 47405, USA}

\author{Yu-Tsung Tai}
\affiliation{Department of Mathematics, Indiana University, Bloomington, Indiana 47405,
USA}
\affiliation{School of Informatics and Computing, Indiana University, Bloomington,
Indiana 47405, USA}

@@

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
In this letter, we extend our previous work on the foundations of computability in quantum mechanics by exploring the application of interval-valued probability measures (IVPM) to achieving a coherent formulation of finite-resource quantum mechanics.

@@

\section{Theory}

\paragraph{Conventional Quantum Theory (\emph{CQT})}

In conventional quantum theory, when we measure a property $\mathbf{O}$
of a physical system whose state is represented by a density matrix~$\rho$,
the probability to obtain the value $\lambda$ of that property is
determined by the Born rule $\mu_{\rho}^{\mathrm{B}}\left(P_{i}\right)=\Tr\left(\rho P_{i}\right)$,
where $P_{i}$ denotes the projector onto the eigenspace corresponding
to the eigenvalue $\lambda_{i}$ of the Hermitian operator $\mathbf{O}$~\cite{Born1983,peres1995quantum,544199,Jaeger2007}.
Alternatively, one may say that the state $\rho$ determines the probability
measure $\mu_{\rho}^{\mathrm{B}}$ defined as a function from the
set $\events$ of all possible projectors $P$ onto the interval $[0,1]$,
and satisfies the following probability axioms~ \cite{10.2307/2308516,gleason1957,Redhead1987-REDINA,Maassen2010}: 
\begin{itemize}
\item $\mu_{\rho}^{\mathrm{B}}(\mathbb{0})=0$. 
\item $\mu_{\rho}^{\mathrm{B}}(\mathbb{1})=1$. 
\item $\mu_{\rho}^{\mathrm{B}}\left(\mathbb{1}-P\right)=
1-\mu_{\rho}^{\mathrm{B}}\left(P\right)$. 
\item For a set of mutually orthogonal projectors $\left\{ P_{i}\right\} _{i=1}^{N}$,
$P_iP_j=\delta_{ij}P_i$,
we have 
\begin{equation}
\mu_{\rho}^{\mathrm{B}}\left(\sum_{i=1}^{N}P_{i}\right)=
\sum_{i=1}^{N}\mu_{\rho}^{\mathrm{B}}\left(P_{i}\right)
\textrm{ .}\label{eq:QuantumProbability-Addition}
\end{equation}
\end{itemize}
Conversely, given any quantum probability measure~$\mu:\events\rightarrow[0,1]$
satisfying the above axioms, Gleason's theorem asserts that there
exists a unique state~$\rho$ such that $\mu=\mu_{\rho}^{\mathrm{B}}$
in a Hilbert space of dimension $d\geq3$~\cite{gleason1957,Redhead1987-REDINA,peres1995quantum}.

@@

\paragraph{Quantum Interval-valued Probability Measures (\emph{IVPMs})}

To formalize the idea of finite precision measurement, we next introduce
the notion of quantum IVPM~$\bar{\mu}$. The latter replaces the
infinitely precise interval $[0,1]$ by a finite collection $\mathscr{I}$
of intervals. Since the intervals $\imposs=\left[0,0\right]$ and
$\necess=\left[1,1\right]$ should be included in $\mathscr{I}$,
and arithmetic operations on intervals satisfy \begin{subequations}\label{eq:interval-operations}
\begin{eqnarray}
 &  & [l_{1},r_{1}]+[l_{2},r_{2}]=[l_{1}+l_{2},r_{1}+r_{2}]\textrm{ and}\\
 &  & x[l,r]=\begin{cases}
[xl,xr] & \textrm{for }x\ge0\textrm{ ;}\\{}
[xr,xl] & \textrm{for }x\le0\textrm{ .}
\end{cases}
\end{eqnarray}
\end{subequations} $\bar{\mu}:\events\rightarrow\mathscr{I}$ is
defined by the axioms: 
\begin{itemize}
\item $\bar{\mu}(\mathbb{0})=\left[0,0\right]$. 
\item $\bar{\mu}(\mathbb{1})=\left[1,1\right]$. 
\item $\bar{\mu}\left(\mathbb{1}-P\right)=\left[1,1\right]-\bar{\mu}\left(P\right)$. 
\item For a set of mutually orthogonal projections $\left\{ P_{i}\right\} _{i=1}^{N}$,
$P_{i}P_{j}=\delta_{ij}P_{i}$, we have 
\begin{equation}
\bar{\mu}\left(\sum_{i=1}^{N}P_{i}\right)\subseteq\sum_{i=1}^{N}\bar{\mu}\left(P_{i}\right)\label{eq:QuantumInterval-valuedProbability-Inclusion}
\end{equation}
\end{itemize}
condition that provides only a bound and not an equivalence as in
Eq.~\eqref{eq:QuantumProbability-Addition}. For example, if we consider
$\left\{ P_{1},P_{2}\right\} =\left\{ P,\mathbb{1}-P\right\} $ and
$\bar{\mu}\left(P\right)=\left[l,r\right]$, we have, when $r-l=\ell>0$,
\[
\bar{\mu}(\mathbb{1})=\left[1,1\right]\subsetneq\left[1-\ell,1+\ell\right]=\bar{\mu}\left(\mathbb{1}-P\right)+\bar{\mu}\left(P\right).
\]

@@

\paragraph{Quantum and Classical IVPMs}

@@

\paragraph{Expectation Values for Quantum IVPMs}

@@

\section{Results}

\paragraph{Gleason's Theorem}

@@

\paragraph{The Kochen-Specker Theorem}

@@

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%

\bibliography{prop}

\end{document}

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%

>>>>

\title{Interval-Valued Quantum Probabilities: Contextuality and the Born Rule}

\author{Yu-Tsung Tai}
\affiliation{Department of Mathematics, Indiana University, Bloomington, Indiana 47405,
USA}
\affiliation{Department of Computer Science, Indiana University, Bloomington,
Indiana 47405, USA}

\author{Andrew J Hanson}
\affiliation{Department of Informatics, Indiana University, Bloomington,
Indiana 47405, USA}

@@

\author{Amr Sabry}
\affiliation{Department of Computer Science, Indiana University, Bloomington,
Indiana 47405, USA}

@@

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\section{Introduction}

In this letter, we extend our previous work on the foundations of
computability in quantum mechanics by exploring the application of
interval-valued probability measures (IVPM) to achieving a coherent
formulation of finite-resource quantum mechanics.

@@

% Textbook quantum mechanics asserts that the probability of a
% measurement can be computed from wave function of the system and the
% measuring observable \emph{exactly} while real measurements must be
% finite precision.  Whether finite precision measurement in reality
% affect the idealized quantum concept, e.g.,
% contextuality~\cite{PhysRevLett.83.3751,Mermin1999,BarrettKent2004},
% and the very foundation of quantum mechanics is still a problem. In
% this letter, we model the finite precision quantum measurement by
% quantum interval-valued probability measures (IVPMs) which provides
% not only a uniform framework to understand Gleason's
% theorem~\cite{gleason1957,Redhead1987-REDINA,peres1995quantum} and the
% Kochen-Specker
% theorem~\cite{kochenspecker1967,peres1995quantum,Redhead1987-REDINA,Griffiths2003}
% but also some examples which cannot be explained by the Born
% rule~\cite{Born1983,peres1995quantum,544199,Jaeger2007}.

% The tension between finite precision measurement and the foundation of
% quantum mechanics sharpened, when the original definition of
% contextuality based on infinitely precise numbers is proved not
% experimental verifiable~\cite{PhysRevLett.83.3751,Kent1999}.  After a
% long debate~\cite{Mermin1999,Peres2003,BarrettKent2004}, people are
% still trying to update the definition of contextuality when they want
% to perform further
% experiments~\cite{Spekkens2005,GuehneKleinmannCabelloEtAl2010,MazurekPuseyKunjwalEtAl2016}.

% We found that finite precision measurement affects not only
% contextuality but also the fundamental rule computing probability from
% a state.  In the early formulation of quantum mechanics, von Neumann
% essentially formulated a quantum state as a density matrix while
% Mackey formulated a quantum state as a quantum probability measure
% which are proved by Gleason to be the essentially same as a density
% matrix~\cite{Varadarajan2008}.  Combining quantum probability measures
% and the classical IVPM~\cite{JamisonLodwick2004} which could be used
% to formulate classical finite precision measurement, we can define
% quantum IVPMs. Surprisingly, there is a quantum IVPM which cannot be
% compatible with any density matrix according to the Born rule. Similar
% to revising the definition of contextuality because of finite
% precision measurement, the very basic definition of a quantum state or
% how to compute the probability from a state might need to be refined
% when we consider the finite precision measurement. This can also
% provide a new quantum phenomena different from the classical world,
% because IVPMs become classical so that the non-Born IVPMs disappears
% when we restricting the domain of quantum IVPMs on a commuting
% sub-event space.

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\section{Quantum Measurements}

A \emph{probability space} is a mathematical abstraction specifying
the necessary conditions for reasoning coherently about collections of
uncertain events~\cite{Kolmogorov1950}.  In the quantum case, the set
of events is the set of \emph{projection operators} $P$. This set
includes the empty projection~$\mathbb{0}$, the identity
projection~$\mathbb{1}$, projections of the form $\proj{\psi}$ where
$\ket{\psi}$ is a pure quantum state, sums of \emph{orthogonal}
projections, and products of \emph{commuting} projections. In a
quantum probability
space~\cite{10.2307/2308516,gleason1957,Redhead1987-REDINA,Maassen2010},
each event $P_i$ is associated with a probability $\mu(P_i)$ subject
to the following axioms: $\mu(\mathbb{0})=0$, $\mu(\mathbb{1})=1$,
$\mu\left(\mathbb{1}-P\right)=1-\mu\left(P\right)$, and for a set of
mutually orthogonal projectors $\left\{ P_{i}\right\} _{i=1}^{N}$, we
have:
\begin{equation}
\mu\left(\sum_{i=1}^{N}P_{i}\right)=
\sum_{i=1}^{N}\mu\left(P_{i}\right)
\textrm{ .}\label{eq:QuantumProbability-Addition}
\end{equation}

\newpage

Conversely, given any quantum probability measure~$\mu:\events\rightarrow[0,1]$
satisfying the above axioms, Gleason's theorem asserts that there
exists a unique state~$\rho$ such that $\mu=\mu_{\rho}^{\mathrm{B}}$
in a Hilbert space of dimension $d\geq3$~\cite{gleason1957,Redhead1987-REDINA,peres1995quantum}.

In conventional quantum theory (CQT), when we measure a property
$\mathbf{O}$ of a physical system whose state is represented by a
density matrix~$\rho$, the probability to obtain the value $\lambda$
of that property is determined by the Born rule
$\mu_{\rho}^{\mathrm{B}}\left(P_{i}\right)=\Tr\left(\rho
  P_{i}\right)$, where $P_{i}$ denotes the projector onto the
eigenspace corresponding to the eigenvalue $\lambda_{i}$ of the
Hermitian operator
$\mathbf{O}$~\cite{Born1983,peres1995quantum,544199,Jaeger2007}.
Alternatively, one may say that the state $\rho$ determines the
probability measure $\mu_{\rho}^{\mathrm{B}}$ defined as a function
from the set $\events$ of all possible projectors $P$ onto the
interval $[0,1]$, and satisfies the following probability axioms~
\cite{10.2307/2308516,gleason1957,Redhead1987-REDINA,Maassen2010}:
\begin{itemize}
\item $\mu_{\rho}^{\mathrm{B}}(\mathbb{0})=0$. 
\item $\mu_{\rho}^{\mathrm{B}}(\mathbb{1})=1$. 
\item $\mu_{\rho}^{\mathrm{B}}\left(\mathbb{1}-P\right)=
1-\mu_{\rho}^{\mathrm{B}}\left(P\right)$. 
\item For a set of mutually orthogonal projectors $\left\{ P_{i}\right\} _{i=1}^{N}$,
$P_iP_j=\delta_{ij}P_i$,
we have 
\begin{equation}
\mu_{\rho}^{\mathrm{B}}\left(\sum_{i=1}^{N}P_{i}\right)=
\sum_{i=1}^{N}\mu_{\rho}^{\mathrm{B}}\left(P_{i}\right)
\textrm{ .}\label{eq:QuantumProbability-Addition}
\end{equation}
\end{itemize}
Conversely, given any quantum probability measure~$\mu:\events\rightarrow[0,1]$
satisfying the above axioms, Gleason's theorem asserts that there
exists a unique state~$\rho$ such that $\mu=\mu_{\rho}^{\mathrm{B}}$
in a Hilbert space of dimension $d\geq3$~\cite{gleason1957,Redhead1987-REDINA,peres1995quantum}.

@@

To formalize the idea of finite precision measurement, we next
introduce the notion of quantum interval-valued probability measures
(\emph{IVPMs})~$\bar{\mu}$. The latter replaces the infinitely precise
interval $[0,1]$ by a finite collection $\mathscr{I}$ of
intervals. Since the intervals $\imposs=\left[0,0\right]$ and
$\necess=\left[1,1\right]$ should be included in $\mathscr{I}$, and
arithmetic operations on intervals
satisfy \begin{subequations}\label{eq:interval-operations}
\begin{eqnarray}
 &  & [l_{1},r_{1}]+[l_{2},r_{2}]=[l_{1}+l_{2},r_{1}+r_{2}]\textrm{ and}\\
 &  & x[l,r]=\begin{cases}
[xl,xr] & \textrm{for }x\ge0\textrm{ ;}\\{}
[xr,xl] & \textrm{for }x\le0\textrm{ .}
\end{cases}
\end{eqnarray}
\end{subequations} $\bar{\mu}:\events\rightarrow\mathscr{I}$ is
defined by the axioms: 
\begin{itemize}
\item $\bar{\mu}(\mathbb{0})=\left[0,0\right]$. 
\item $\bar{\mu}(\mathbb{1})=\left[1,1\right]$. 
\item $\bar{\mu}\left(\mathbb{1}-P\right)=\left[1,1\right]-\bar{\mu}\left(P\right)$. 
\item For a set of mutually orthogonal projections $\left\{ P_{i}\right\} _{i=1}^{N}$,
$P_{i}P_{j}=\delta_{ij}P_{i}$, we have 
\begin{equation}
\bar{\mu}\left(\sum_{i=1}^{N}P_{i}\right)\subseteq\sum_{i=1}^{N}\bar{\mu}\left(P_{i}\right)\label{eq:QuantumInterval-valuedProbability-Inclusion}
\end{equation}
\end{itemize}
condition that provides only a bound and not an equivalence as in
Eq.~\eqref{eq:QuantumProbability-Addition}. For example, if we consider
$\left\{ P_{1},P_{2}\right\} =\left\{ P,\mathbb{1}-P\right\} $ and
$\bar{\mu}\left(P\right)=\left[l,r\right]$, we have, when $r-l=\ell>0$,
\[
\bar{\mu}(\mathbb{1})=\left[1,1\right]\subsetneq\left[1-\ell,1+\ell\right]=\bar{\mu}\left(\mathbb{1}-P\right)+\bar{\mu}\left(P\right).
\]

@@

@@

@@

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\section{Quantum States}

%%%%%
\subsection{Gleason}

@@

%%%
\subsection{Kochen-Specker}

@@

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\section{Conclusion}

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\bibliography{prop}
\end{document}
